---
title: "Data Visualization"
output: html_document
date: "2023-11-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 01 Data Visualization Basics

From: [Posit Primers](https://posit.cloud/learn/primers/1.1)

### 01.01 Import Data

```{r}
library(ggplot2)
mpg
```

### 01.02 Make first plot

```{r}
ggplot(data=mpg)+ geom_point(mapping=aes(x=displ,y=hwy))
```

**Explanation**:\
1. ``` ggplot(``) ```: Creates coordinate system that can add layers to, First argument is dataset, by itself creates empty graph

2.   `geom_point()`:
    1.  Adds layer of points to the empty plot
    2.  \``mapping`\`-argument: Defines which variables mapped to hich axes, always *paired* with function `aes()` which use gather all mappings

**Summary**:

1.  Start graph with `ggplot()`
2.  Add elements to graph with a `geom_` function
3.  Select variables with the `mapping=aes()` argument

**Common mistakes**:

1.  Adding the `+` at the wrong place, has to come at end of line

### 01.03 Aesthetics

**What is an aesthetic**:

-   Visual property of objects in plot (`size, shape, color`)

-   Use *level* to describe aesthetic properties

-   To use: Set name of aesthetic (aka. `color` or `shape` equal to name of variable inside `mapping=aes()`

-   Find out aesthetics availabele for `geom`: `?geom_line`

```{r}
ggplot(data=mpg)+ geom_point(mapping=aes(x=displ, y=hwy, color=class))
```

#### 01.03.01 The alpha aesthetic

**What is the alpha aesthetic**

-   controls the transparency

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha=class))

```

#### 01.03.02 Aesthetics for continuos variables

```{r}
# Map cty to color
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color=cty))
```

### 01.03.03 Using multiple aesthetics

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color=class, size=class, shape=class))
```

#### 01.03.04 Map aesthetic to something other than variable name

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color=displ<5))
```

### 01.03.05 Set all to same thing

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")

```

**MERKE**: Settiing the aesthetic fo a value like `blue` above: outside of `aes()` , Mapping the aesthetic to a **varibale**: Inside of `aes()`

## 01.04 Geometric Objects

**What is a geom**

-   Geom: geometrical object that a plot uses to represent observations (bar chart use bar geoms, boxplots use boxplot geoms, scatterplots use **point** geoms)

-   To change. chage the `geom` function that was added to `ggplot()`

-   Overview can be found: <https://exts.ggplot2.tidyverse.org/gallery/>

-   Learn about geoms: `?geom_smooth`

-   

**Examples of geoms**

```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy),se=TRUE)

```

```{r}
?geom_smooth
```

-   `se`: Can be `TRUE` or `FALSE`, show standard error ribbon
